---
title: "Vis I"
author: "Yifei Yu"
date: "2024-10-04"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

Import the weather data

```{r}
weather_df = 
  read_csv("weather_df.csv") |> 
  filter(
    id %in% c("USW00094728", "USW00022534", "USS0023B17S"), 
    
    date >= "2021-01-01", 
    date <= "2022-12-31"
  ) |> 
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin,
    tmax = tmax) |>
  select(name, id, date, prcp, tmin, tmax) 
```


Making our first plot.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

Check why some rows are missing.

```{r}
weather_df |> 
  filter(is.na(tmax))
##is this missing
```



## Fancier scatterplots!

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)

### adding alpha is to make the plot more transparent
### size makes the points look smaller
### three smooth curves
```


Where you define aesthetics can matter

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)

### Only one smooth curve
```


Use faceting real quick

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)

### . means everything
### facet_grid(name ~ .) reverse the graph
```


Let's make a somewhat more interesting scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


## Learning Assessment

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(
    tmax_fahr = tmax*(9/5) +32 ,
    tmin_fahr = tmin*(9/5) +32
  ) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE)
```


## Small things

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
### density plot
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "blue")
```


## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

How would i fix this? maybe facet?

```{r}
weather_df |> 
  ggplot(aes( x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```


maybe a density plot?

```{r}
weather_df |> 
  ggplot(aes( x = tmin, fill = name)) +
  geom_density(alpha = .3)

## alphs is the transparent level
```

```{r}
weather_df |> 
  ggplot(aes( x = name, y = tmin)) +
  geom_boxplot()
```


violin plots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```


ridge plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Learning Assessment -- univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```


```{r}
weather_df |> 
  ggplot(aes(x = name, y = prcp)) +
  geom_boxplot()
```


```{r}
weather_df |> 
  filter(prcp > 10, prcp < 1000) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```




## Saving and embedding plots

saving plots

```{r}
ggp_weather =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

ggsave("plots/ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```


embedding plots

```{r, fig.width = 12, fig.asp = .6}
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

```


